Trade Dependence

### 2022-12-12 article/article.qmd
Update scope description

The very broad measure of trade dependence dominant in the literature examines the total trade between states $i$ and $j$ relative to state $i$'s gross domestic product (GDP) in year $t$:

$$
TD_{ij,t} = \frac{M_{ij,t} + X_{ij,t}}{GDP_{i,t}}
$$

Some scholars argue that this measure captures the cost state $i$ considers when deciding whether to adopt some unfavorable foreign policy towards state $j$ [@oneal1997]. I disagree.

@@

@@

This analysis focuses on all directed state dyads in all years between 2000 and 2014. This includes `r nrow(scope) |> comma()` observations.

@@

My dependent variable focuses on the onset of Militarized Interstate Disputes, or events in which one or more states threatens, displays, or uses force against another state or states. I include two binary dependent variables. First, I identify all dyad-years in which any MID was initiated. Second, I identify all dyad-years in which any MID involving at least one battle-related death was initiated. These variables are coded 1 if state $i$ initiated the MID against state $j$ in a given year. Otherwise, they receive a value of 0.

>>>>

@@

Returning to our motivating question: does trade dependence deter states from initiating militarized conflict against each other? The commercial peace theory argues that trade dependence increases the expected cost of initiating conflict against a trade partner, decreasing the expected net benefits derived from such an action and, therefore, reducing the likelihood a state will initiate conflict. This leads to the following testable hypothesis:

> H1: As product-level trade dependence between two states increases, the likelihood that a state will initiate militarized conflict against its trade partner decreases.

@@

This analysis focuses on all directed state dyads in 2014. This includes `r nrow(scope) |> comma()` observations. We are interested in whether or not a trade dependent state initiates conflict against another state. This is a directed question. We need to know whether the potential initiator is dependent on their opponent, not simply whether either partner has a dependence on their other.

The narrow timeframe included in this analysis is driven by data constraints. I only have access to complete trade and militarized interstate dispute data for 2014. A natural first step for extending this analysis is including a broader timeframe in the analysis.

@@

My dependent variable focuses on the onset of militarized interstate disputes, or events in which one or more states threaten, display, or use force against another state or states. I include two binary dependent variables. First, I identify all dyad-years in which any MID was initiated. Second, I identify all dyad-years in which any MID involving at least one battle-related death was initiated. These variables are coded 1 if state $i$ initiated the MID against state $j$ in a given year. Otherwise, they receive a value of 0.
